Verifying Spatial Properties of Stencil Computations

### 2016-11-16 paper.tex
rev of the eval

We have processed nearly one million lines of Fortran code using our
CamFort tool and examined the output to try and determine how useful
and applicable it may be. The body of Fortran code is collected from
many computational science projects for which we believe our tool will
be potentially helpful.

@@

\subsection{Core hypotheses}

\paragraph{Hypothesis 1: Stencil computations
(as in \Cref{def:stencil}) are a common programming pattern
in numerical code}

%    numStencilSpecs: 60525
% 4338 files, linesTotal: 1357377, linesParsed: 959427
% Number of stencils (tickAssign) / number of loops
% Percentage of a code base that is taken up by stencils

Out of 1.35 million lines of code that we fed into our program, and
the 959,427 lines that we were able to parse, we identified 60,525
cases that fit our criteria for stencil specification.

\paragraph{Hypothesis 2: Most stencil computations
have a regular pattern which can be captured with a simple
language of contiguous regions}

@@

In the analysed body of code we found that we were able to deduce
some form of stencil specification from approximately 30\% of the
potential stencils examined.

%
\subsection{Analysing the design of our specifications}

@@

55,504 specifications involved only reflexive regions, of which 15,823
only had specifications for some of the dimensions, while 39,681 were
reflexive on all dimensions.

@@

% singleAction: 4837
% singleActionIrr: 1532
A single-action specification consists of no more than one forward,
backward, or centered region and can be combined with any number of
reflexive regions. We identified 4,837 single-action specifications,
of which 1,532 were single-action with an \texttt{irreflexive}
modifier.

@@

    % multiAction: 156
    % multiActionMulOnly: 146
A multi-action specification consists of at least two uses of forward,
backward, or centered regions, and can be combined with any number of
reflexive regions. We identified 156 multi-action specifications out
of which 146 had regions combined only with $\term{*}$.

@@

Bounded specifications were fairly unusual, with only 157
\texttt{atMost} and 36 \texttt{atLeast}, the latter of which were
always also bounded on both sides.
    % atLeast: 36
    % atMost: 157
    % boundedBoth: 36

>>>>

%We have processed nearly one million lines of Fortran code using our
%CamFort tool and examined the output to determine how useful
%and applicable it may be. Our corpus of Fortran code is collected from
%many computational science projects for which we believe our tool will
%be potentially helpful.

@@

We first examined how frequently stencil computations occur in our
corpus of Fortran code. This was carried out by running CamFort in
whole-code-base inference mode. We parsed 959,427 lines of Fortran
code and found that 10\% (97,439) of statements have a left-hand side
is an array subscript on neighbourhood indices. This supports the idea
that stencil-like computations are common in scientific code.

We would not expect to infer a stencil for each of these statements
because our stencil language places significant restrictions on the
types of array-subscript-statements that we will class as a
stencil. In fact, CamFort was able to infer a stencil from 30\% of
these statements. A single statement can produce multiple stencils and
we ended up with 60,525 stencils. This shows that we can still express
a large number of stencil uses within our high-level abstraction.

%    numStencilSpecs: 60525
% 4338 files, linesTotal: 1357377, linesParsed: 959427
% Number of stencils (tickAssign) / number of loops
% Percentage of a code base that is taken up by stencils
%Out of 1.35 million lines of code that we fed into our program, and
%the 959,427 lines that we were able to parse, we identified 60,525
%cases that fit our criteria for stencil specification.

%\paragraph{Hypothesis 2: Most stencil computations
%have a regular pattern which can be captured with a simple
%language of contiguous regions}

@@

%In the analysed body of code we found that we were able to deduce
%some form of stencil specification from approximately 30\% of the
%potential stencils examined.

%
%\subsection{Analysing the design of our specifications}

@@

The vast majority of stencils we found were relatively simple but we
found significant numbers of more complex shapes. We grouped common
patterns in to categories which we describe below.

\textbf{All pointed} 55,504 of the stencils we found involved 
only pointed regions. 39,681 of these were pointed in all
dimensions. Common examples of this were pointwise transformations on
data (such as scaling).

\textbf{Single action} specifications consist of no more than one 
forward, backward, or centered region and can be combined with any
number of pointed regions. We identified 4,837 single-action
specifications, of which 1,532 were single-action with a
\texttt{nonpointed} modifier.

\textbf{Multi-action} specifications consist of at least two uses 
of forward, backward, or centered regions, and can be combined with
any number of pointed regions. We identified 156 multi-action
specifications out of which 146 had regions combined only with
$\term{*}$.

\textbf{Bounded} specifications were fairly unusual, with only 157
\texttt{atMost} and 36 \texttt{atLeast}, the latter of which were
always also bounded on both sides.

@@

% singleAction: 4837
% singleActionIrr: 1532

@@

    % multiAction: 156
    % multiActionMulOnly: 146

@@

    % atLeast: 36
    % atMost: 157
    % boundedBoth: 36
